Privacy-Preserving Localisation Using Private Linear-Combination Aggregation

### 2020-11-10 main.tex
Full read-through with fixes updates and some missing notational consistency

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Distributed state estimation and localisation methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor or navigator information during localisation protocols, thus not preserving participants’ data privacy. Existing approaches which provide such guarantees fail to address sensor and navigator privacy in some common non-linear measurement model applications and forfeit broad applicability. We define the novel notion of linear-combination aggregation encryption and provide a cryptographically secure instance that is applied to the Extended Information Filter with range-sensor measurements. Our method keeps navigator location, sensor locations and sensor measurements private during navigation, and is implemented and run on simulations to evaluate its accuracy and performance. The novel computationally plausible and provably secure range-based localisation filter has direct applications to environments where nodes are not trusted or data is considered sensitive.
\end{abstract}

@@

The remainder of the paper is structured as follows. In Section \ref{sec:problem_statement} we introduce the localisation problem, the relevant aggregation opertation for its solution and the restrictions on involved parties. Sections \ref{sec:crypto_prelim} and \ref{sec:lcao_definition} will summarise cryptographic preliminaries and introduce a novel security notion for our required aggregation before defining and proving an encryption scheme which satisfies it. Sections \ref{sec:loc_prelim} and \ref{sec:priv_localisation} introduce localisation preliminaries and our privacy-preserving localisation method using the private aggregation scheme. We show and discuss simulation results in Section \ref{sec:sim_and_results} and conclude our work in Section \ref{sec:conclusion}.

@@

\section{Literature Overview}
In this section, we will discuss methods for providing security guarantees in signal processing tasks, and their applications to localisation and estimation problems.

@@

    \item[Aggregator Oblivious Encryption] This method of encryption formalises the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in these schemes is, however, strictly limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter sum or mean calculation \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.

@@

    \item[Partially Homomorphic Encryption] A simplification of fully homomorphic schemes; these schemes provide only a subset, typically one, of algebraic operations on encrypted data \cite{elgamalPublicKeyCryptosystem1985,paillierPublicKeyCryptosystemsBased1999,bonehEvaluating2DNFFormulas2005}. Reduced computational requirements and implementation simplicity have made partially homomorphic encryption schemes commonly used in privacy-preserving signal processing tasks, including private matrix multiplication \cite{kogisoCyberSecurityEnhancementNetworked2015}, set intersection computation \cite{kerschbaumOutsourcedPrivateSet2012} and control input aggregation \cite{alexandruEncryptedCooperativeControl2019}. However, these works are relatively restricted in application due to the limited operations provided.

@@

    \item[Multi-Client Function Encryption] These schemes allow the computing of a specific function given encrypted inputs \cite{bonehFunctionalEncryptionDefinitions2011,goldwasserMultiinputFunctionalEncryption2014}. It is suitable for data-sensitive distributed processing as encryptions must be combined before a result is obtained, and can be considered a generalisation of the aggregation schemes described earlier. While general schemes that allow computing arbitrary functions are computationally expensive and complicate security definitions in distributed settings \cite{agrawalFunctionalEncryptionNew2013}, partial schemes supporting only some functions, such as vector dot products \cite{chotardDecentralizedMultiClientFunctional2018}, exist with formal security proofs, but have yet to find applications in more complicated signal processing tasks.

@@

Of the aforementioned encryption schemes, some have found uses in privacy-preserving localisation and estimation tasks and we summarise these in the next section.

@@

\subsection{Encrypted Localisation and Estimation}
Localisation methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors. We aim to provide an accurate localisation filter in the presence of range-only sensors.

Relevant estimation methods include both Bayesian estimation methods which make assumptions about target trajectories to form their location estimates, typically based on the Kalman Filter and later derivatives \cite{kalmanNewApproachLinear1960,mutambaraDecentralizedEstimationControl1998}, and time-independent methods where location estimates are made independently of each other \cite{alanwarPrOLocResilientLocalization2017,lazosSeRLocSecureRangeindependent2004}. The work in \cite{alanwarPrOLocResilientLocalization2017} uses partially homomorphic encryption to compute time-independent location estimates while preserving the privacy of sensor measurements and locations. The work proposes two approaches, using either polygon intersections suffering from geometric dilution of precision, or alternating projection requiring repeated interactive comparison protocols. Neither method considers measurement uncertainty, and no estimate errors are provided.

Bayesian approaches can achieve better accuracy due to their exploitation of process dynamic knowledge. In \cite{aristovEncryptedMultisensorInformation2018}, partially homomorphic encryption is used to encrypt measurement information and produce location estimates at a navigator. While producing estimate errors and requiring only uni-directional communication, this method exclusively supports linear measurement models and is therefore not suitable for range-only measurements. Encrypted model predictive control methods \cite{farokhiPrivacyDynamicalSystems2020} also introduce some privacy-preserving methods relevant to estimation. \cite{alexandruEncryptedCooperativeControl2019} proposes secure distributed control input aggregation using a novel weighted sum aggregation scheme, but requires repeated key distribution among all involved parties. In \cite{farokhiSecurePrivateControl2017}, a secure cloud-based control aggregator using partially homomorphic encryption is proposed, but is suitable only for untrusted service-providing cloud architectures.

While we have focused on the application of encryption schemes to localisation algorithms, homomorphic encryption schemes also require suitable number encoding when used with real numbers. As is the case with localisation sensors, real number measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. Google's Encrypted BigQuery Client \cite{googleEncryptedbigqueryclient2015} provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In \cite{farokhiSecurePrivateControl2017}, a simpler alternative for encoding is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings. We will discuss and justify our choice of encoding scheme when introducing the localisation filter.

@@

\section{Problem Statement} \label{sec:problem_statement}
The localisation scenario we consider in this work is that of model-based self-navigation using range-only sensors. We consider localisation in the two-dimensional case for simplicity but will derive methods suitable for an extension to the three-dimensional equivalent.

The navigator state is defined as 
\begin{equation}
    \rvec{x} = 
    \begin{bmatrix}
        \rv{x} & \rv{dx} & \rv{y} & \rv{dy}
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $\rvec{w}_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependant on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$, and measurement function $h_i$ for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
defined as
\begin{equation}
    \begin{split}
        h_i(\rvec{x}) &= \left\lVert
        \begin{bmatrix}
            \rv{x} & \rv{y}
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(\rv{x}-s_{x,i})^2 + (\rv{y}-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn state estimates of $\rvec{x}_{k}$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$, measurement variances $r_i$ or their measurements $\rv{z}_{k,i}$ at any time $k$. We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. However, the additional safety to passengers provided by accurate aircraft localisation, may be a goal of all those involved justifying their cooperation.

@@

To support multiple aggregations at each time-step $k$, as is required for the aggregation of multi-dimensional data, aggregation instances have been denoted by $t$. The desired bi-direction communication protocol at instance $t$ consists of of the computation of linear combinations of navigator weights by sensors, before their privacy-preserving aggregation by the navigator. That is, $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ are broadcast by the navigator, linear combinations $l^{(t)}_i=\sum^m_{j=1}a_{j,i}^{(t)}\omega_i^{(t)}$ are computed by each sensor $1\leq i\leq n$, and aggregation is computed back at the navigator. This has been summarised in Figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\vspace{\baselineskip}
\begin{tikzpicture}[font=\footnotesize]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and combination step};
    % Navigator
    \fill (3.25,4.5) [blue!40] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.6725) {plane};
    % Sensors
    \node at (1.5,0.375) {Sensor $1$};
    \fill [red!40] (0.25,0.625) rectangle (2.875,1.625);
    \node at (1.625,1.125) {$\displaystyle l_1^{(t)} \!=\! \sum^m_{j=1}a_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,0.375) {Sensor $n$};
    \fill [red!40] (3.625,0.625) rectangle (6.25,1.625);
    \node at (5,1.125) {$\displaystyle l_n^{(t)} \!=\! \sum^m_{j=1}a_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,1.125) circle (0.05);
    \fill [black] (3,1.125) circle (0.05);
    \fill [black] (3.25,1.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4) (5,1.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4) (1.5,1.875)};
    \fill [lightgray] (2.125,3.375) rectangle (4.375,2.625);
    \node at (3.25,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

@@

In Section \ref{sec:priv_localisation} we will show how this protocol is sufficient to compute the required estimation information at the sensors required to update the navigator's estimation filter.

\begin{remark}
    From the aggregation definition above and the distributive property of real numbers, it is clear that an alternative method to the bi-directional communication requirement would be computing $\sum^{m}_{j=1}\omega_j^{(t)}\sum^{n}_{i=1} a_{i,j}^{(t)}$ by the navigator, requiring only uni-directional communication by sending values $a_{i,j}^{(t)}, 1 \leq j \leq m$ from each sensor $i$. We justify the use of bi-directional communication by reducing communication costs when the number of weights is larger than the number of sensors, $m>n$, and by sending fewer weights in the presence of repeats, as will be shown to be the case in Section \ref{subsec:algorithm}.
\end{remark}

@@

\subsection{Participant Capabilities} \label{subsec:capabilities}
Before we can define concrete security requirements and a filtering protocol, we require some assumptions on the capabilities of the navigator and sensors in our problem scenario.
\begin{LaTeXdescription}
    \item[Global navigator broadcast] From the decomposition defined above, we know that bi-directional communication between navigator and sensors is required. We make the assumption that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.

@@

\section{Cryptography Preliminaries} \label{sec:crypto_prelim}
When defining our aggregation security requirements and encryption scheme, we will reference some existing cryptographic security notions, the additively homomorphic Paillier encryption scheme and the Joye-Libert private aggregation scheme.

@@

Encryption of plaintext $a^{(t)}_{i} \in \mathbb{Z}_N$ to ciphertext $c^{(t)}_{i} \in \mathbb{Z}_{N^2}$ at instance $t$ is computed by user $i$ as
\begin{equation}
    c^{(t)}_{i} = (N+1)^{a^{(t)}_{i}} H(t)^{sk_i} \pmod{N^2}\,,
\end{equation}
where the $H(t)^{sk_i}$ can be considered the noise term which hides the again easily computable discrete logarithm $g^{a^{(t)}_{i}} \pmod{N^2}$, where $g=N+1$.

@@

    \item[$\mathsf{Enc}(pk_0, x)$] Encryption of values is computed as a Paillier encryption with implicit generator $g=N+1$. This is given by
    \begin{equation}
        \mathcal{E}_{pk_0}(a) = (N+1)^{x}\rho^N \pmod{N^2}\,, \label{eqn:our_scheme_encrypt}
    \end{equation}
    for a randomly chosen $\rho \in \mathbb{Z}_N$.

@@

We define a subset of rational numbers in terms of a range $M \in \mathbb{N}$ and fractional precision $\phi \in \mathbb{N}$. This contrasts with the common definition given in terms of total bits and fractional bits \cite{oberstarFixedPointRepresentationFractional2007,schulzedarupEncryptedCooperativeControl2019,farokhiSecurePrivateControl2017}, but allows for a direct mapping to integer ranges which are not powers of two. Rational subset $\mathbb{Q}_{M,\phi}$ is given by
\begin{equation}
    \mathbb{Q}_{M,\phi} = \left\{q \,\middle|\, \phi q \in \mathbb{N} \wedge -\left\lfloor\frac{M}{2}\right\rfloor \leq q < \left\lfloor\frac{M}{2}\right\rfloor \right\}\,,
\end{equation}
and we quantise any real number $a$ by taking the nearest rational $q \in \mathbb{Q}_{M,\phi}$. That is, $\argmin_{q\in\mathbb{Q}_{M,\phi}} |a-q|$. In this form, mapping rationals $\mathbb{Q}_{M,\phi}$ to the encryption scheme range $\mathbb{Z}_N$ is achieved by choosing $M=N$, and handling negatives with modulo arithmetic. In addition, we note that the Q number format requires a precision factor $\phi$ to be removed after each encoded multiplication, which is not supported by our encryption scheme. This is captured by a third parameter $d$; the number of multiplication factors to add or remove from encodings.

The function for combined quantisation and encoding $\mathsf{E}_{M,\phi,d}(a)$, of a given a real number $a \in \mathbb{R}$, integer range $\mathbb{Z}_M$, and the desired scaling for $d$ prior encoded multiplications, is given by
\begin{equation}
    \mathsf{E}_{M,\phi,d}(a) = \left\lfloor \phi^{d+1} a \right\rceil \pmod{M}\,. \label{eqn:encode}
\end{equation}
Decoding of an integer $u \in \mathbb{Z}_M$, is given by
\begin{equation}
    \mathsf{E}^{-1}_{M,\phi,d}(u) \!=\! 
    \begin{dcases}
        \frac{u\hspace{-8pt}\pmod{M}}{\phi^{d+1}}, &u\hspace{-8pt}\pmod{M} \leq \left\lfloor\frac{M}{2}\right\rfloor \\
        -\frac{M - u\hspace{-8pt}\pmod{M}}{\phi^{d+1}}, &\text{otherwise} \\
    \end{dcases}.\label{eqn:decode}
\end{equation}

@@

The choice of a high fractional precision $\phi$ may reduce quantisation errors introduced in \eqref{eqn:encode}, however, risks an overflow following too many multiplications. Given the largest number of expected multiplications $d_{max}$, and the largest expected decoded value $a$, the parameter should be chosen such that the following condition holds
\begin{equation}
    \left|\phi^{d_{max}+1}a\right| < \left\lfloor \frac{M}{2} \right\rfloor\,.
\end{equation}
In practice, $N$ is typically very large ($N>2^{1024}$) and this condition can be ignored when $M=N$.

@@

Algorithms \ref{alg:nav_prediction}, \ref{alg:measurement_info} and \ref{alg:nav_update} have been summarised graphically in Figure \ref{fig:alg_steps}, where encryptions and encodings of vectors and matrices have been denoted with $\mathcal{E}_{pk_0,sk_i}(\cdot)$ and $\mathsf{E}_m(\cdot)$ for brevity, and represent element-wise operations with the same parameters.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Distributed state estimation and localisation methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor or navigator information during localisation protocols, thus not preserving participants’ data privacy. Existing approaches which provide such guarantees fail to address sensor and navigator privacy in some common non-linear measurement model applications and forfeit broad applicability. We define the novel notion of linear-combination aggregation encryption and provide a cryptographically secure instance that is applied to the Extended Information Filter with range-sensor measurements. Our method keeps navigator location, sensor locations and sensor measurements private during navigation, and has been implemented to evaluate its accuracy and performance. The novel computationally plausible and provably secure range-based localisation filter has direct applications to environments where participants are not trusted or data is considered sensitive.
\end{abstract}

@@

The remainder of the paper is structured as follows. In Section \ref{sec:lit_overview} we give an overview of the relevant literature to the topic and in Section \ref{sec:problem_statement} we introduce the localisation problem, the relevant aggregation opertation for its solution and the restrictions on involved parties. Sections \ref{sec:crypto_prelim} and \ref{sec:lcao_definition} will summarise cryptographic preliminaries and introduce a novel security notion for our required aggregation before defining and proving an encryption scheme which satisfies it. Sections \ref{sec:loc_prelim} and \ref{sec:priv_localisation} introduce localisation preliminaries and our privacy-preserving localisation method using the private aggregation scheme. We show and discuss simulation results in Section \ref{sec:sim_and_results} and conclude our work in Section \ref{sec:conclusion}.

@@

\section{Literature Overview} \label{sec:lit_overview}
In this section, we will discuss methods for providing security guarantees in signal processing tasks, and their applications to localisation and estimation problems.

@@

    \item[Aggregator Oblivious Encryption] This method of encryption formalises the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in these schemes is, however, strictly limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter sums or mean calculations \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.

@@

    \item[Partially Homomorphic Encryption] A simplification of fully homomorphic schemes; these schemes provide only a subset of, typically one, algebraic operations on encrypted data \cite{elgamalPublicKeyCryptosystem1985,paillierPublicKeyCryptosystemsBased1999,bonehEvaluating2DNFFormulas2005}. Reduced computational requirements and implementation simplicity have made partially homomorphic encryption schemes commonly used in privacy-preserving signal processing tasks, including private matrix multiplication \cite{kogisoCyberSecurityEnhancementNetworked2015}, set intersection computation \cite{kerschbaumOutsourcedPrivateSet2012} and control input aggregation \cite{alexandruEncryptedCooperativeControl2019}. However, as with Aggregator Oblivious Encryption, these works are relatively restricted in application due to the limited operations provided.

@@

    \item[Multi-Client Function Encryption] These schemes allow the computing of a specific function given encrypted inputs \cite{bonehFunctionalEncryptionDefinitions2011,goldwasserMultiinputFunctionalEncryption2014}. They are suitable for data-sensitive distributed processing as encryptions must be combined before a result is obtained, and can be considered a generalisation of the aggregation schemes described earlier. While general schemes that allow computing arbitrary functions are computationally expensive and complicate security definitions in distributed settings \cite{agrawalFunctionalEncryptionNew2013}, partial schemes supporting only some functions, such as vector dot products \cite{chotardDecentralizedMultiClientFunctional2018}, exist with formal security proofs, but have yet to find applications in more complicated signal processing tasks.

@@

In the context of privacy-preserving localisation and state estimation, some of the aforementioned schemes have found applications. These have been summarised in the next section.

@@

\subsection{Encrypted Localisation and Estimation}
Localisation methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors. We aim to provide a computationally plausible and accurate localisation filter in the presence of range-only sensors while providing uncertainties associated with estimates.

Relevant estimation methods include both Bayesian estimation methods which make assumptions about target trajectories to form their location estimates, typically based on the Kalman Filter and later derivatives \cite{kalmanNewApproachLinear1960,mutambaraDecentralizedEstimationControl1998}, and time-independent methods where location estimates are made independently of each other \cite{alanwarPrOLocResilientLocalization2017,lazosSeRLocSecureRangeindependent2004}. The work in \cite{alanwarPrOLocResilientLocalization2017} uses partially homomorphic encryption to compute time-independent location estimates while preserving the privacy of sensor measurements and locations. The work proposes two approaches, using either polygon intersections suffering from geometric dilution of precision, or alternating projection requiring repeated interactive comparison protocols. Neither method considers measurement uncertainty, and no estimate errors are provided, making them unsuitable for our desired solution.

Bayesian approaches can achieve better accuracy due to their exploitation of process dynamic knowledge. In \cite{aristovEncryptedMultisensorInformation2018}, partially homomorphic encryption is used to encrypt measurement information and produce location estimates at a navigator. While producing estimate errors and requiring only uni-directional communication, this method exclusively supports linear measurement models and is therefore not suitable for range-only measurements. Encrypted model predictive control methods \cite{farokhiPrivacyDynamicalSystems2020} also introduce some privacy-preserving methods relevant to estimation. \cite{alexandruEncryptedCooperativeControl2019} proposes secure distributed control input aggregation using a novel weighted sum aggregation scheme, but requires a repeated, slow, key distribution among all involved parties. In \cite{farokhiSecurePrivateControl2017}, a secure cloud-based control aggregator using partially homomorphic encryption is proposed, but is suitable only for untrusted service-providing cloud architectures and handles only the aggregation step homomorphically, making its application to localisation impractical.

Above we have focused on the application of encryption schemes to localisation-relevant algorithms, however, homomorphic encryption schemes also require suitable number encoding when used with real numbers. As is the case with range sensors, real-valued measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. Google's Encrypted BigQuery Client \cite{googleEncryptedbigqueryclient2015} provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In \cite{farokhiSecurePrivateControl2017}, a simpler alternative for encoding is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings. We will discuss and justify our choice of encoding scheme when introducing the localisation filter.

From the descriptions above, we can see that no existing methods provide a general-purpose privacy-preserving localisation method supporting range-only sensors. In the next section, we will formalise our problem and its accompanying security requirements.

@@

\section{Problem Statement} \label{sec:problem_statement}
The localisation scenario we consider in this work is that of Bayesian self-navigation using range-only sensors. We consider localisation in the two-dimensional case for simplicity but will derive methods suitable for an extension to the three-dimensional equivalent.

The navigator state is defined as 
\begin{equation}
    \rvec{x} = 
    \begin{bmatrix}
        \rv{x} & \rv{dx} & \rv{y} & \rv{dy}
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $\rvec{w}_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependant on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$ and measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
defined as
\begin{equation}
    \begin{split}
        h_i(\rvec{x}) &= \left\lVert
        \begin{bmatrix}
            \rv{x} & \rv{y}
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(\rv{x}-s_{x,i})^2 + (\rv{y}-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn state estimates of $\rvec{x}_{k}$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$, measurement variances $r_i$ or their measurements $\rv{z}_{k,i}$ at any time $k$. We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localisation, however, may be a goal of all those involved, justifying their cooperation.

@@

To support multiple aggregations at each time-step $k$, as is required for the aggregation of multi-dimensional data, aggregation instances have been denoted by a different \textit{instance} variable $t$. The desired bi-direction communication protocol at instance $t$ consists of of the computation of linear combinations of navigator weights by sensors, before their privacy-preserving aggregation by the navigator. That is, $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ are broadcast by the navigator, linear combinations $l^{(t)}_i=\sum^m_{j=1}a_{j,i}^{(t)}\omega_i^{(t)}$ are computed by each sensor $1\leq i\leq n$, and aggregation is computed back at the navigator. This has been summarised in Figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\vspace{\baselineskip}
\begin{tikzpicture}[font=\footnotesize]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and combination step};
    % Navigator
    \fill (3.25,4.5) [blue!40] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.6725) {plane};
    % Sensors
    \node at (1.5,0.375) {Sensor $1$};
    \fill [red!40] (0.25,0.625) rectangle (2.875,1.625);
    \node at (1.625,1.125) {$\displaystyle l_1^{(t)} \!=\! \sum^m_{j=1}a_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,0.375) {Sensor $n$};
    \fill [red!40] (3.625,0.625) rectangle (6.25,1.625);
    \node at (5,1.125) {$\displaystyle l_n^{(t)} \!=\! \sum^m_{j=1}a_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,1.125) circle (0.05);
    \fill [black] (3,1.125) circle (0.05);
    \fill [black] (3.25,1.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4) (5,1.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4) (1.5,1.875)};
    \fill [lightgray] (2.125,3.375) rectangle (4.375,2.625);
    \node at (3.25,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

@@

In Section \ref{sec:priv_localisation} we will show how this protocol is sufficient to compute the required estimation information at the sensors needed to update the navigator's estimation filter.

\begin{remark}
    From the aggregation definition above and the distributive property of real numbers, it is clear that an alternative method to the bi-directional communication requirement would be computing $\sum^{m}_{j=1}(\omega_j^{(t)}\sum^{n}_{i=1} a_{i,j}^{(t)})$ by the navigator, requiring only uni-directional communication by sending values $a_{i,j}^{(t)}, 1 \leq j \leq m$ from each sensor $i$. We justify the use of bi-directional communication by reducing communication costs when the number of weights is larger than the number of sensors, $m>n$, and by sending fewer weights in the presence of repeats, as will be shown to be the case in Section \ref{subsec:algorithm}.
\end{remark}

@@

\subsection{Participant Capabilities} \label{subsec:capabilities}
Before we can define formal security requirements and a filtering protocol, we require some assumptions on the capabilities of the navigator and sensors in our problem scenario.
\begin{LaTeXdescription}
    \item[Global navigator broadcast] From the decomposition defined above, we know that bi-directional communication between navigator and sensors is required. We make the assumption that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.

@@

\section{Cryptography Preliminaries} \label{sec:crypto_prelim}
When defining our aggregation security requirements and encryption scheme we will reference some existing cryptographic security notions, the additively homomorphic Paillier encryption scheme and the Joye-Libert private aggregation scheme.

@@

Encryption of plaintext $a^{(t)}_{i} \in \mathbb{Z}_N$ to ciphertext $c^{(t)}_{i} \in \mathbb{Z}_{N^2}$ at instance $t$ is computed by user $i$ as
\begin{equation}
    c^{(t)}_{i} = (N+1)^{a^{(t)}_{i}} H(t)^{sk_i} \pmod{N^2}\,,
\end{equation}
where $H(t)^{sk_i}$ can be considered the noise term which hides the again easily computable discrete logarithm $g^{a^{(t)}_{i}} \pmod{N^2}$, where $g=N+1$.

@@

    \item[$\mathsf{Enc}(pk_0, x)$] Encryption of values is computed as a Paillier encryption with implicit generator $g=N+1$. This is given by
    \begin{equation}
        \mathcal{E}_{pk_0}(x) = (N+1)^{x}\rho^N \pmod{N^2}\,, \label{eqn:our_scheme_encrypt}
    \end{equation}
    for a randomly chosen $\rho \in \mathbb{Z}_N$.

@@

We define a subset of rational numbers in terms of a range $M \in \mathbb{N}$ and fractional precision $\phi \in \mathbb{N}$. This contrasts with the common definition given in terms of total bits and fractional bits \cite{oberstarFixedPointRepresentationFractional2007,schulzedarupEncryptedCooperativeControl2019,farokhiSecurePrivateControl2017}, but allows for a direct mapping to integer ranges which are not powers of two. Rational subset $\mathbb{Q}_{M,\phi}$ is given by
\begin{equation}
    \mathbb{Q}_{M,\phi} = \left\{q \,\middle|\, \phi q \in \mathbb{N} \wedge -\left\lfloor\frac{M}{2}\right\rfloor \leq \phi q < \left\lfloor\frac{M}{2}\right\rfloor \right\}\,,
\end{equation}
and we quantise any real number $a$ by taking the nearest rational $q \in \mathbb{Q}_{M,\phi}$. That is, $\argmin_{q\in\mathbb{Q}_{M,\phi}} |a-q|$. In this form, mapping rationals $\mathbb{Q}_{M,\phi}$ to the encryption scheme range $\mathbb{Z}_N$ is achieved by choosing $M=N$, and handling negatives with modulo arithmetic. In addition, we note that the Q number format requires a precision factor $\phi$ to be removed after each encoded multiplication, which is not supported by our encryption scheme. This is captured by a third parameter $d$; the number of multiplication factors to add or remove from encodings.

The function for combined quantisation and encoding $\mathsf{E}_{M,\phi,d}(a)$, of a given real number $a \in \mathbb{R}$, integer range $\mathbb{Z}_M$, and the desired scaling for $d$ prior encoded multiplications, is given by
\begin{equation}
    \mathsf{E}_{M,\phi,d}(a) = \left\lfloor \phi^{d+1} a \right\rceil \pmod{M}\,. \label{eqn:encode}
\end{equation}
Decoding of an integer $u \in \mathbb{Z}_M$, is given by
\begin{equation}
    \mathsf{E}^{-1}_{M,\phi,d}(u) \!=\! 
    \begin{dcases}
        \frac{u\hspace{-8pt}\pmod{M}}{\phi^{d+1}}, &u\hspace{-8pt}\pmod{M} \leq \left\lfloor\frac{M}{2}\right\rfloor \\
        -\frac{M - u\hspace{-8pt}\pmod{M}}{\phi^{d+1}}, &\text{otherwise} \\
    \end{dcases}.\label{eqn:decode}
\end{equation}

@@

The choice of a high fractional precision $\phi$ may reduce quantisation errors introduced in \eqref{eqn:encode}, however, risks an overflow following too many multiplications. Given the largest number of expected multiplications $d_{max}$, and the largest expected decoded value $a_{max}$, the parameter should be chosen such that the following condition holds
\begin{equation}
    \left|\phi^{d_{max}+1}a_{max}\right| < \left\lfloor \frac{M}{2} \right\rfloor\,.
\end{equation}
In practice, $N$ is typically very large ($N>2^{1024}$) and this condition can be ignored when $M=N$.

@@

Algorithms \ref{alg:nav_prediction}, \ref{alg:measurement_info} and \ref{alg:nav_update} have been summarised graphically in Figure \ref{fig:alg_steps}, where encryptions and encodings of vectors and matrices have been denoted with $\mathcal{E}_{pk_0,sk_i}(\cdot)$ and $\mathsf{E}_d(\cdot)$ for brevity, and represent element-wise operations with the same parameters.
